---
title: "Untitled"
author: "Elsie Horne"
date: "02/03/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
```

## R Markdown

```{r}
library(tidyverse)
library(glue)
```

```{r paths}
# path for released results
release_folder <- here::here("release20220226")
```

```{r read-meta-data}
# read study parameters
study_parameters <- readr::read_rds(
  here::here("output", "lib", "study_parameters.rds"))

# read subgroups
subgroups <- readr::read_rds(
  here::here("output", "lib", "subgroups.rds"))
subgroup_labels <- seq_along(subgroups)
subgroup_order <- c(4,1,3,2)
subgroups_long <- if_else(
    subgroups %in% subgroups[c(2,3)],
    as.character(glue("{subgroups} and not clinically vulnerable")),
    subgroups
  )

# outcomes 
outcomes <- readr::read_rds(
  here::here("output", "lib", "outcomes.rds"))
outcomes_order <- c(which(outcomes == "covidadmitted"),
                    which(outcomes == "coviddeath"),
                    which(outcomes == "postest"),
                    which(outcomes == "noncoviddeath"),
                    which(outcomes == "anytest"))
select_outcomes <- unname(outcomes[outcomes_order])

```


```{r}
# read estimates data
estimates_all <- readr::read_csv(
  here::here(release_folder, "estimates_all.csv")) %>%
  mutate(across(model, ~as.integer(str_remove(.x, "unadjusted")))) %>%
  mutate(across(c(estimate, conf.low, conf.high), 
                ~format(round(.x, digits=2), nsmall=2, trim = TRUE))) %>%
  mutate(value = glue("{estimate} (95% CI {conf.low}-{conf.high})"))

# read metaregression results
metareg_results_k <- readr::read_rds(
  here::here(release_folder, "metareg_results_k.rds")) %>%
  mutate(across(subgroup, as.integer)) %>%
  mutate(across(outcome,
                factor,
                levels = unname(outcomes[outcomes_order]),
                labels = names(outcomes[outcomes_order]))) %>%
      mutate(across(comparison,
                  factor,
                  levels = c("BNT162b2", "ChAdOx", "both"),
                  labels = c("BNT162b2 vs unvaccinated", "ChAdOx vs unvaccinated", "BNT162b2 vs ChAdOx"))) %>%
   mutate(across(c(logrhr, logrhr_lower, logrhr_higher), 
                ~format(round(exp(.x), digits=2), nsmall=2, trim = TRUE))) %>%
  select(subgroup, comparison, outcome, k, rhr = logrhr, rhr_lower = logrhr_lower, rhr_higher = logrhr_higher)
```


```{r}
aHRs <- local({
  
  data <- estimates_all %>% 
    filter(
      model ==2,
      variable == "k",
      !reference_row
    )
  
  data <- data %>%
    group_split(subgroup) %>%
    as.list()
  
  names(data) <- c("16-64", "18-39", "40-64", "65+")
  
  for (i in seq_along(data)) {
    
    comparisons <- sort(unique(data[[i]]$comparison))
    
    data[[i]] <- data[[i]] %>%
      group_split(comparison) %>%
      as.list()
    
    names(data[[i]]) <- comparisons
    
    for (j in seq_along(data[[i]])) {
      
      outcomes <- sort(unique(data[[i]][[j]]$outcome))
      
      data[[i]][[j]] <- data[[i]][[j]] %>%
        group_split(outcome) %>%
        as.list()
      
      names(data[[i]][[j]]) <- outcomes
      
      for (k in seq_along(data[[i]][[j]])) {
        data[[i]][[j]][[k]] <- data[[i]][[j]][[k]] %>%
          arrange(label) %>%
          select(value) %>%
          unlist() %>%
          unname()
      }
      
    }
    
  }
  
  return(data)
  
})
```


```{r}
ratios <- local({
  
  data <- metareg_results_k %>%
  select(-k) %>%
  distinct(subgroup, comparison, outcome, .keep_all = TRUE) %>%
  mutate(across(comparison,
                factor,
                levels = c("BNT162b2 vs unvaccinated",
                           "ChAdOx vs unvaccinated",
                           "BNT162b2 vs ChAdOx"),
                labels = c("BNT162b2", "ChAdOx", "both"))) %>%
  mutate(across(outcome,
                factor,
                levels = names(outcomes),
                labels = unname(outcomes))) %>%
    mutate(value = glue("{rhr_higher} (95% CI {rhr_lower}-{rhr_higher})"))
  
  data <- data %>%
    group_split(subgroup) %>%
    as.list()
  
  names(data) <- c("16-64", "18-39", "40-64", "65+")
  
  for (i in seq_along(data)) {
    
    comparisons <- sort(unique(data[[i]]$comparison))
    
    data[[i]] <- data[[i]] %>%
      group_split(comparison) %>%
      as.list()
    
    names(data[[i]]) <- comparisons
    
    for (j in seq_along(data[[i]])) {
      
      outcomes <- sort(unique(data[[i]][[j]]$outcome))
      
      data[[i]][[j]] <- data[[i]][[j]] %>%
        group_split(outcome) %>%
        as.list()
      
      names(data[[i]][[j]]) <- outcomes
      
      for (k in seq_along(data[[i]][[j]])) {
        data[[i]][[j]][[k]] <- data[[i]][[j]][[k]] %>%
          select(value) %>%
          unlist() %>%
          unname()
      }
      
    }
    
  }
  
  return(data)
  
})

```


For COVID-19 hospital admission, estimated aHRs comparing BNT162b2 and ChAdOx with unvaccinated in the 65+ subgroup were `r aHRs[["65+"]][["BNT162b2"]][["covidadmitted"]][1]` and `r aHRs[["65+"]][["ChAdOx"]][["covidadmitted"]][1]` respectively at 3-6 weeks, waning to `r aHRs[["65+"]][["BNT162b2"]][["covidadmitted"]][6]` and `r aHRs[["65+"]][["ChAdOx"]][["covidadmitted"]][6]` respectively at 23-26 weeks. 
The ratios of aHRs were `r ratios[["65+"]][["BNT162b2"]][["covidadmitted"]]` and `r ratios[["65+"]][["ChAdOx"]][["covidadmitted"]]` per period respectively. 
In the 16-64 subgroup the equivalent aHRs were `r aHRs[["16-64"]][["BNT162b2"]][["covidadmitted"]][1]` and `r aHRs[["16-64"]][["ChAdOx"]][["covidadmitted"]][1]` for BNT162b2 and ChAdOx respectively at 3-6 weeks, waning to `r aHRs[["16-64"]][["BNT162b2"]][["covidadmitted"]][6]` and `r aHRs[["16-64"]][["ChAdOx"]][["covidadmitted"]][6]` respectively at 23-26 weeks.
The ratios of aHRs were `r ratios[["16-64"]][["BNT162b2"]][["covidadmitted"]]` and `r ratios[["16-64"]][["ChAdOx"]][["covidadmitted"]]` per period respectively. 
