---
title: "Investigating the distribution of positive tests around COVID-19 hospitalisations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r libraries}
library(tidyverse)
library(lubridate)
library(kableExtra)
library(glue)
```

```{r functions}
source(here::here("analysis", "functions", "data_process_functions.R"))
```

Our plan is to replace the outcome variable `covidadmitted`, which currently corresponds to the date of hospitalisation with a COVID-19 code in any field.
The new variable will correspond to "an episode of COVID-19 resulting in hospitalisation", and will take the date of the recorded evidence of the episode of COVID-19 (evidence being a positive COVID-19 test or COVID-19 related hospital admission).
We have carried out some exploratory data analysis to inform the algorithm that we will use to derive this new variable.

First, Table 1 shows the number of eligible patients who have a COVID-19 hospitalisation in their record, and whether this was recorded in the primary field.
It is possible that someone would have a hospital admission with COVID-19 in a non-primary field, then at a later date have a COVID-19 hospital admission with COVID-19 in a primary field.
Of the patients with COVID-19 hospital admission in a primary field, Table 2 shows the proportion for whom this occurred on the same date as a COVID-19 hospital admission in any field.

```{r load-data}
# read data for ever covariates
data_ever <- arrow::read_feather(
  file = here::here("output", "input_ever.feather")) 

# clean data
data_0 <- data_ever %>%
  select(patient_id, covidadmitted_post_date, covidadmittedprimary_post_date, starts_with("covidadmitted_postest")) %>%
  mutate(across(contains("_date"), 
                ~ floor_date(
                  as.Date(.x, format="%Y-%m-%d"),
                  unit = "days"))) 
```

#### Table 1. Patients with a COVID-19 hospital admission by recording field
```{r table1}
data_0 %>%
  mutate(across(c(covidadmitted_post_date, covidadmittedprimary_post_date),
         ~!is.na(.x))) %>%
  group_by(covidadmitted_post_date, covidadmittedprimary_post_date) %>%
  count() %>%
  ungroup() %>%
  mutate(percent = round(100*n/sum(n), 2)) %>%
  mutate(across(n, scales::comma, accuracy=1)) %>%
  kbl(
    col.names = c("any", "primary", "n", "percent")
  ) %>%
  kable_styling(full_width = F) %>%
  add_header_above(c("field" = 2, " " = 2))
```

#### Table 2. Hospitalistion with COVID in primary and any field on same date
```{r table2}
data_0 %>%
  filter(
    !is.na(covidadmitted_post_date),
    !is.na(covidadmittedprimary_post_date)
    ) %>%
  mutate(
    same_date = covidadmitted_post_date == covidadmittedprimary_post_date
    ) %>%
  group_by(same_date) %>%
  count() %>%
  ungroup() %>%
  mutate(percent = round(100*n/sum(n), 2)) %>%
  mutate(across(n, scales::comma, accuracy=1)) %>%
  kbl(
    col.names = c("Same date", "n", "percent")
  ) %>%
  kable_styling(full_width = F)
```


```{r data1}
data_1 <- data_0 %>%
  filter(!is.na(covidadmitted_post_date)) %>%
  pivot_longer(
    cols = c(covidadmitted_post_date, covidadmittedprimary_post_date),
    values_drop_na = TRUE
  ) %>%
  mutate(
    covidadmitted_field = if_else(
      name %in% "covidadmittedprimary_post_date",
      "primary",
      "any")
    ) %>%
  select(-name) %>%
  rename(covidadmitted_date = value) %>%
  pivot_longer(
    cols = starts_with("covidadmitted_postest"),
    names_pattern = "covidadmitted_postest_(.)_date",
    names_transform = as.integer,
    values_drop_na = TRUE
  ) %>%
  select(-name) %>%
  mutate(diff = as.numeric(value - covidadmitted_date)) %>%
  filter(-42 <= diff, diff <= 14) %>%
  group_by(patient_id, covidadmitted_field) %>%
  arrange(value, .by_group = TRUE) %>%
  mutate(
    sequence = row_number(),
    min_postest_date = min(sequence),
    max_postest_date = max(sequence)
    ) %>%
  ungroup()
```


For the patients with a COVID-19 hospitalisation, we counted the number of positive tests in their record during the window 6 weeks before and 2 weeks after the date of their COVID-19 hospitalisation. 
To speed up computation, the number of positive tests was capped at 6.
These results (split by field) are given in Table 3.

#### Table 3. Number of positive tests per patient during the window around hospitalisation date
```{r table3}
n_tests <- local({
  counts <- data_1 %>% 
    distinct(patient_id, covidadmitted_field, max_postest_date) %>%
    group_by(max_postest_date, covidadmitted_field) %>%
    count() %>%
    ungroup() %>%
    mutate(across(c(max_postest_date, n), as.integer))
  
  total_any <- sum(counts[counts$covidadmitted_field == "any",]$n)
  total_primary <- sum(counts[counts$covidadmitted_field == "primary",]$n)
  
  counts %>%
    pivot_wider(
      names_from = covidadmitted_field,
      values_from = n
    ) %>%
    rename(n_postests = max_postest_date) %>%
    add_row(
      n_postests = 0L, 
      any = nrow(data_0 %>% filter(!is.na(covidadmitted_post_date))) - total_any,
      primary = nrow(data_0 %>% filter(!is.na(covidadmittedprimary_post_date))) - total_primary
    ) %>%
    arrange(n_postests) %>%
    mutate(across(c(any, primary), 
                  ~if_else(is.na(.x),0L,.x))) %>%
    mutate(across(c(any, primary), 
                  ~str_c(scales::comma(.x, accuracy = 1), " (", round(100*.x/sum(.x), 2), ")")))
  
})

n_tests %>%
  kbl() %>%
  kable_styling(full_width = F) %>%
  add_header_above(c(" " = 1, "field" = 2))
```


```{r myhist-function}
my_hist <- function(data, filename) {
  
  plot <- data %>%
    mutate(across(covidadmitted_field,
                  ~str_c("field = ", .x))) %>%
  ggplot(aes(x = diff, fill = covidadmitted_field)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
    stat_count(
      alpha = 0.8
    ) +
    facet_wrap(~covidadmitted_field, ncol=1, scales = "free_y") +
  scale_x_continuous(
    breaks = seq(-42,14,7),
    limits = c(-42,14),
    labels = as.character(seq(-6,2,1))
  ) +
  labs(
    x = "Weeks since COVID-19 hospital admission",
    y = "Number of positive tests"
  ) +
  theme_bw() +
  theme(
    axis.title.x = element_text(margin = margin(t=10)),
    axis.title.y = element_text(margin = margin(r=10)),
    legend.position = "none"
  ) +
    coord_cartesian(ylim = c(6,NA))
  
# ggsave(
#   plot = plot,
#   filename = here::here("output", "eda", glue("{filename}.png"))
# )

return(plot)
  
}
```

We used histograms to investigate the distribution of positive test date relative to COVID-19 hospitalisation date.
Figure 1 shows the distribution of positive test dates for all tests during the window, and Figure 2 shows the distribution of each patients earliest positive test date during the window.
The y-axes start at 6 to mask bars that correspond to <=5 patients.

#### Figure 1. Distribution of all positive test dates within window relative to date of hospitalisation
```{r figure1}
# distribution of all positive test dates relative to date of hospital admission
data_1 %>%
  my_hist(filename = "covidadmitted_postest_all")
```

#### Figure 2. Distribution of earliest positive test dates within window relative to date of hospitalisation
```{r figure2}
# distribution of all positive test dates relative to date of hospital admission
data_1 %>%
  filter(sequence == min_postest_date) %>%
  my_hist(filename = "covidadmitted_postest_min")
```

Although most patients tested positive within two weeks of the hospital admission, some tested positive >2 weeks before.
To investiagte this, we restricted the sample to those who tested positive 4-6 weeks before their admission date.
Of these patients, we counted how many also had a positive test 0-4 weeks before their admission date (see Table 4).
Figure 3 shows the distribution of positive tests in the window 0-4 weeks before and 0-2 weeks after the hospital admission date for patients with a positive test 4-6 weeks before admission date.

#### Table 4. Number of patients with a positive test 4-6 weeks before hospitalisation and during the window 4 weeks before to 2 weeks after hospitalisation
```{r table4}
postest_after_4wks <- data_1 %>%
  filter(diff < -28) %>%
  select(patient_id, covidadmitted_field) %>%
  left_join(data_1, by = c("patient_id", "covidadmitted_field")) %>%
  mutate(postest_after_4wks = diff > -28)

postest_after_4wks %>%
  distinct(patient_id, covidadmitted_field, postest_after_4wks) %>%
  group_by(covidadmitted_field, postest_after_4wks) %>%
  count() %>%
  ungroup() %>%
  pivot_wider(
    names_from = covidadmitted_field,
    values_from = n
  ) %>%
  mutate(across(c(any, primary), ~if_else(is.na(.x), 0L, .x))) %>%
  mutate(across(c(any, primary), 
                  ~str_c(scales::comma(.x, accuracy = 1), " (", round(100*.x/sum(.x), 2), ")"))) %>%
  kbl() %>%
  kable_styling(full_width = F) %>%
  add_header_above(c(" " = 1, "field" = 2))
```


#### Figure 3. Distribution of all positive test dates within window relative to date of hospital admission for those who have a positive test 4-6 weeks before the date of hospitalisation 
```{r figure3}
# distribution of all positive test dates relative to date of hospital admission
postest_after_4wks %>%
  filter(diff > -28) %>%
  my_hist(filename = "covidadmitted_postest_min")
```

